Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-29 paper2_JNLS/paper.tex
NM

\begin{abstract} 
Recent laboratory experiments of \boetal revealed the emergence of anomalous statistics in a randomized wave-field that encounters an abrupt change in bottom topography. \maetal established a theoretical foundation, based on analysis of the truncated Korteweg–de Vries (TKdV) system, that successfully captures several key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the depth change, and is corroborated by numerical simulations of the deterministic dynamics. 
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments and a quantitative comparison can be made. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in a spectacular fashion by new experimental measurements.
\end{abstract}

>>>>

\begin{abstract} 
\boetal experimentally demonstrated that an abrupt change in bottom topography can trigger anomalous statistics in a randomized field of surface waves. Motivated by these observations, \maetal developed a theoretical framework, based on deterministic and statistical analysis of the truncated KdV system, that successfully captures key qualitative features of the experiments, including the robust emergence of anomalous statistics in the outgoing waves. Here, we extend these parallel experimental and modeling efforts with several new findings that have resulted from synergetic interaction between the two. By fleshing out details of the non-dimensionalization, we precisely relate model parameters to physical ones, thereby enabling calibration of the model inverse temperature to the specific conditions present in the experiments and, ultimately, a quantitative comparison between theory and experiments. We find the theoretically predicted distributions of surface displacement to match the experimental measurements with surprising quantitive detail. Prompted by the presence of surface slope in the TKdV Hamiltonian, we experimentally examine surface slope statistics and compare them to theoretical predictions. Both show significant spread of the slope distributions with long, flat tails in the outgoing wave field. Finally, the theory predicts a peculiar relationship between the outgoing displacement skewness and the change in slope variance, specifically how their ratio depends on the wave amplitude and depth ratio. New experimental measurements confirm this prediction in spectacular fashion.
\end{abstract}
